---
title: "Analysis of the Music of Smog and Townes Van Zandt"
author: "Harpeth Lee"
date: "4/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(rnoaa)
library(rvest)
library(httr)
library(RColorBrewer)
library(spotifyr)
library(glue)
library(stringr)
library(tidytext)
library(ggrepel)
library(viridis)
library(wesanderson)
library(wordcloud)
library(genius)
library(RColorBrewer)
```


## Analysis of Spotify Data

```{r, eval = FALSE}
Sys.setenv(SPOTIFY_CLIENT_ID = 'e2749fe5f78d409e9ca00307f79d25cf')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '925cd5121e004fba8491aeb9bb7c14c5')

access_token <- get_spotify_access_token()
```

```{r, eval = FALSE}
## Creates data frames with artist info
townes_raw <- get_artist_audio_features('townes van zandt')
smog_raw <- get_artist_audio_features('smog')
```

```{r, eval = FALSE}
## Keeps only variables that I think might be interesting to work with
townes_raw <- townes_raw %>%
  select(c(artist_name, track_name, album_name, album_release_year, danceability, speechiness, energy, loudness, liveness, valence, acousticness, instrumentalness, key_mode, tempo, time_signature, duration_ms)) %>%
  filter(album_name != "Be Here To Love Me (Soundtrack)")

smog_raw <- smog_raw %>%
  select(c(artist_name, track_name, album_name, album_release_year, danceability, speechiness, energy, loudness, liveness, valence, acousticness, instrumentalness, key_mode, tempo, time_signature, duration_ms))
```

```{r, eval = FALSE}
write_csv(townes_raw, "townes.csv")
write_csv(smog_raw, "smog.csv")
```

```{r}
townes <- read_csv("townes.csv")
smog <- read_csv("smog.csv")
```


```{r}
## Ranks both artists albums by how sad they are (valence is a measurement created by spotify to measure how positive a track is), includes other stats.

townes %>%
  group_by(album_name) %>%
  summarise(valence = mean(valence), acousticness = mean(acousticness), energy = mean(energy), danceability = mean(danceability)) %>%
  arrange(valence)

smog %>%
  group_by(album_name) %>%
  summarise(valence = mean(valence), acousticness = mean(acousticness), energy = mean(energy), danceability = mean(danceability)) %>%
  arrange(valence)
```


```{r}
## Compares the distribution of track danceability scores for both artists

ggplot() +
  geom_density(data = smog, mapping = aes(x = danceability), color = "blue") +
  geom_density(data = townes, mapping = aes(x = danceability), color = "red") +
  labs(x = "Danceability", y = "Density",
       title = "Is Smog or Townes Van Zandt's music more danceable?",
       subtitle = "blue line corresponds to Smog, red to Townes Van Zandt") +
  theme_minimal()
```

I decided to analyze the music of my two most listened to artists, Smog and Townes Van Zandt. I found that Smog makes sadder sounding music based on the average valences of the two's albums as Smog's "happiest" album still had a lower average valence than Van Zandt's "saddest". Furthermore, Van Zandt's music is more danceable than Smog's. This suggests that there may be correlation between a song's danceability and valence, a relationship that would make sense. Calculating the correlation between the two, we can see that there is some correlation, but not a strong relationship between the two variables in the data.

## Analysis of Lyrics

```{r}
rbind(smog, townes) %>%
  summarise(correlation = cor(valence, danceability))
```

```{r}
knock_knock <- genius_album(artist = "Smog", album = "Knock Knock")
wild_love <- genius_album(artist = "Smog", album = "Wild Love")
old_quarter <- genius_album(artist = "Townes Van Zandt", album = "Live at the Old Quarter")
high_low <- genius_album(artist = "Townes Van Zandt", album = "High, Low and In Between")
```

```{r}
plot_cloud <- function(x, color = viridis(20, direction = -1)){
  x %>%
    mutate(lyric = tolower(lyric)) %>%
    unnest_tokens(output = word, input = lyric, token = "words") %>%
    anti_join(stop_words, by = "word") %>%
    count(word, sort = TRUE) %>%
    head(20) %>%
    with(wordcloud(word, n, random.order = FALSE, scale = c(4, 1), colors = color))
}
```

```{r}
plot_cloud(knock_knock, color = brewer.pal(10, "Blues"))
plot_cloud(wild_love, color = brewer.pal(10, "Greys"))
plot_cloud(old_quarter, color = brewer.pal(10, "YlGn"))
plot_cloud(high_low, color = brewer.pal(10, "YlOrRd"))
``` 

```{r}
nrc <- get_sentiments("nrc")
afinn <- get_sentiments("afinn")

nrc_table <- function(x){
  x %>%
    unnest_tokens(output = word, input = lyric, token = "words") %>%
    mutate(word = tolower(word)) %>%
    anti_join(stop_words, by = "word") %>%
    left_join(nrc, by = "word") %>%
    filter(!is.na(sentiment)) %>%
    count(sentiment) %>%
    arrange(-n)
}
```

```{r}
nrc_table(knock_knock)
nrc_table(wild_love)
nrc_table(old_quarter)
nrc_table(high_low)
```

```{r}
knock_knock <- knock_knock %>%
  mutate(album = "Knock Knock", artist = "Smog")
wild_love <- mutate(wild_love, album = "Wild Love", artist = "Smog")
old_quarter <- mutate(old_quarter, album = "Live at the Old Quarter", artist = "Townes Van Zandt")
high_low <- mutate(high_low , album = "High, Low and In Between", artist = "Townes Van Zandt")
```

```{r}
knock_knock %>%
  rbind(wild_love) %>%
  rbind(old_quarter) %>%
  rbind(high_low) %>%
  unnest_tokens(output = word, input = lyric, token = "words") %>%
  mutate(word = tolower(word)) %>%
  count(album, artist, word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  left_join(nrc, by = "word") %>%
  filter(!is.na(sentiment)) %>%
  filter(sentiment != "positive" & sentiment != "negative") %>%
  group_by(album) %>%
  mutate(sentiment = fct_reorder(sentiment, n)) %>%
  ggplot(aes(x = sentiment, y = n, fill = album)) +
  geom_col() +
  facet_grid(vars(album)) +
  labs(x = "Sentiment", y = "Frequency") +
  theme_minimal()
```

```{r}
nrc_plot <- function(x, plot_title = NULL, fill_color = "grey2"){
  nrc_table(x) %>%
    filter(sentiment != "positive" & sentiment != "negative") %>%
    mutate(sentiment = fct_reorder(sentiment, n)) %>%
    ggplot(aes(x = sentiment, y = n)) +
    geom_col(fill = fill_color) +
    labs(title = plot_title, x = "Sentiment", y = "Frequency") +
    theme_minimal()
}
```

```{r}
nrc_plot(knock_knock, plot_title = "Knock Knock Sentiment Frequency", "dodgerblue4")
nrc_plot(wild_love, plot_title = "Wild Love Sentiment Frequency", "wheat4")
nrc_plot(old_quarter, "Live at the Old Quarter Sentiment Frequency", "palegreen4")
nrc_plot(high_low, "High, Low and In Between Sentiment Frequency", "orangered4")
```

```{r}
afinn_summary <- function(x){
  x  %>%
    unnest_tokens(output = word, input = lyric, token = "words") %>%
    mutate(word = tolower(word)) %>%
    count(word, sort = TRUE) %>%
    left_join(afinn, by = "word") %>%
    summarise(mean_sentiment = mean(value, na.rm = TRUE), median_sentiment = median(value, na.rm = TRUE))
}
```

```{r}
afinn_impact <- function(x, head = 10) {
  x %>%
    unnest_tokens(output = word, input = lyric, token = "words") %>%
    mutate(word = tolower(word)) %>%
    count(word, sort = TRUE) %>%
    left_join(afinn, by = "word") %>%
    filter(!is.na(value)) %>%
    mutate(impact = n*value) %>%
    arrange(desc(abs(impact))) %>%
    head(head)
}
```


```{r}
afinn_impact(knock_knock)
afinn_impact(wild_love)
afinn_impact(old_quarter)
afinn_impact(high_low)
```

```{r}
impact_plot <- function(x, plot_title = NULL) {
  afinn_impact(x, 15) %>%
    mutate(word = fct_reorder(word, impact)) %>%
    ggplot(aes(x = word, y = impact, fill = as.factor(value))) +
    geom_col() +
    geom_text(aes(label = word), angle = -15) +
    labs(title = plot_title) +
    scale_fill_viridis(discrete = TRUE, begin = 0.5, name = "afinn value") +
    theme_minimal() +   
    theme(axis.title.x = element_blank(), axis.text.x = element_blank())
}
```


```{r}
impact_plot(knock_knock, "High Impact Words on Knock Knock")
impact_plot(wild_love, "High Impact Words on Wild Love")
impact_plot(old_quarter, "High Impact Words on Live at the Old Quarter")
impact_plot(high_low, "High Impact Words on High, Low and In Between")
```

I decided to go back and compare the sentiments on my favorite two albums by each of them. In this previous question, I found that Smog's music was sadder according to spotify's analysis. I wanted to know if his music also contained more a more negative sentiment as well. For both artists, one album was overall more positive while one was more negative. These seemed to average out to be about the same. Ultimately, I did see that sentiment analysis runs into problems with sarcasm and further context. Especially with the album Wild Love, it seemed that a lot of the words that gave it a highly positive sentiment were not so positive when considered in context.

